\documentclass[12pt]{exam}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}

% Header and Footer
\pagestyle{headandfoot}
\firstpageheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 4 of Algebra 2}
\runningheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 4 of Algebra 2}
\firstpagefooter{}{Page \thepage\ of \numpages}{}
\runningfooter{}{Page \thepage\ of \numpages}{}

% Title
\newcommand{\examtitle}{Polynomial Operations and Factoring}

% Instructions
\newcommand{\instructions}{
    \noindent\rule{\textwidth}{0.5pt}
    \begin{center}
    \textbf{Instructions:} Answer all questions to the best of your ability. Show all your work in the space provided for full credit.
    \end{center}
    \noindent\rule{\textwidth}{0.5pt}
}

\begin{document}

\begin{center}
\textbf{\Large \examtitle} \\
\vspace{0.5cm}
\makebox[0.4\textwidth]{Name: \enspace\hrulefill}
\hspace{0.1\textwidth}
\makebox[0.4\textwidth]{Date: \enspace\hrulefill}
\end{center}

\instructions
\vspace{0.5cm}

\begin{questions}

\pointsinrightmargin

\question[10]
Find all solutions to each of the following equations:
\begin{parts}
    \part $r^2 - 7r = 0$
    \vspace*{2cm}
    
    \part $x^2 + 3x = 7x - x^2$
    \vspace*{2cm}
    
    \part $2t^2 = 242$
    \vspace*{2cm}
    
    \part $16 - y^2 = -4$
    \vspace*{2cm}
\end{parts}

\question[8]
Find all solutions to the equation $t^4 - 11t^2 + 18 = 0$.

\textit{Hint:} Let $u = t^2$ and solve for $u$ first.
\vspace*{4cm}

\question[10]
Solve the inequality $m^2 + 6m + 5 \geq 0$, and graph the solutions on the number line.
\vspace*{5cm}

\newpage

\question[8]
For what values of $r$ is $2r^2 - 3r > -7$?
\vspace*{4cm}

\question[12]
Factor $x^4 + 4y^4$ (completely).

\vspace*{5cm}

\question[15]
Factor each of the following:
\begin{parts}
    \part $a^3 + 27$
    \vspace*{2cm}
    
    \part $a^3b^3 + 8c^3$
    \vspace*{2cm}
    
    \part $2r^3 - 16$
    \vspace*{2cm}
    
    \part $1000 - x^6y^3$
    \vspace*{2cm}
\end{parts}

\newpage

\question[12]
\begin{parts}
    \part The expression $x^5 + y^5$ can be written as the product of $x + y$ and another factor. Find that other factor.
    \vspace*{5cm}
    
    \part The expression $x^7 + y^7$ can be written as the product of $x + y$ and another factor. Find that other factor.
    \vspace*{5cm}
    
    \part Write $x^{2n+1} + y^{2n+1}$ as the product of two factors.
    \vspace*{5cm}
    
    \part Why does the factorization in the previous part fail when the powers of $x$ and $y$ are even? In other words, why can we not factor $x^4 + y^4$ or $x^6 + y^6$ using the patterns we found in the first three parts?
    \vspace*{3cm}
\end{parts}

\newpage

\question[10]
Solve each polynomial equation by factoring:
\begin{parts}
    \part $x^3 - 8 = 0$
    \vspace*{2cm}
    
    \part $x^3 + 64 = 0$
    \vspace*{2cm}
    
    \part $2x^3 - 16x = 0$
    \vspace*{2cm}
    
    \part $x^4 - 16 = 0$
    \vspace*{2cm}
\end{parts}

\question[8]
Find the GCF of $54x^7t^3$, $90x^5t^2$, and $108x^4t$.
\begin{parts}
    \part $18x^7t$
    \part $9x^2t$
    \part $540x^4t^3$
    \part $36xt$
\end{parts}
\vspace*{4cm}

\newpage

\question[12]
A rectangle is twice as long as it is wide. If its length is increased by 4 cm and its width is decreased by 3 cm, the new rectangle formed has an area of 100 cm². Find the dimensions of the original rectangle.
\vspace*{6cm}

\end{questions}

\end{document} 