\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{parskip}

\begin{document}

Owen Payton \\
B.S. Candidate in Mathematics, 2022--2026 \\
University of California, Santa Barbara \\
\texttt{owenpayton@ucsb.edu} \\
August 22, 2025

\vspace{1em}

Dear Admissions Committee,

I am pleased to recommend Krish Arora, a rising ninth grader who completed an intensive, self-directed Algebra II summer program under my guidance. The curriculum was intentionally more rigorous and open-ended than a standard Algebra II course. My approach emphasized developing a higher-mathematical perspective on core algebraic ideas. For example, beyond routine factoring, we examined polynomials over different fields ($\mathbb{R}[x]$, $\mathbb{C}[x]$) and discussed how foundational number-theoretic results for integers---such as the division algorithm---extend to polynomials.

Krish was assessed on both proof-based and computational problems, with many drawn from competitions such as the AMC 10/12 and HMMT. He consistently absorbed new concepts quickly, connected them to prior knowledge, and asked original, incisive follow-up questions---clear indicators of mathematical talent and curiosity.

Based on his performance, I am confident that, after an introductory course in proofs, Krish could hold his own in upper-division undergraduate courses such as number theory or linear algebra. He has demonstrated complete command of all thirteen chapters of \textit{Algebra and Trigonometry: Structure and Method, Book 2}.

I strongly recommend Krish for admission. Please feel free to contact me at the email above if you have any questions.

Sincerely,

Owen Payton

\end{document}


