\documentclass[12pt]{exam}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}

% Header and Footer
\pagestyle{headandfoot}
\firstpageheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 10 of Algebra 2}
\runningheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 10 of Algebra 2}
\firstpagefooter{}{Page \thepage\ of \numpages}{}
\runningfooter{}{Page \thepage\ of \numpages}{}

% Title
\newcommand{\examtitle}{Exponential and Logarithmic Functions}

% Instructions
\newcommand{\instructions}{
    \noindent\rule{\textwidth}{0.5pt}
    \begin{center}
    \textbf{Instructions:} Answer all questions to the best of your ability. Show all your work in the space provided for full credit.
    \end{center}
    \noindent\rule{\textwidth}{0.5pt}
}

\begin{document}

\begin{center}
\textbf{\Large \examtitle} \\
\vspace{0.5cm}
\makebox[0.4\textwidth]{Name: \enspace\hrulefill}
\hspace{0.1\textwidth}
\makebox[0.4\textwidth]{Date: \enspace\hrulefill}
\end{center}

\instructions
\vspace{0.5cm}

\begin{questions}

\pointsinrightmargin

\question[8]
Find the base $n$ such that $\log_n 4\sqrt{2} = 10$.
\vspace*{4cm}

\question[10]
Find $x$ if $\log_9(2x - 7) = \frac{3}{2}$.
\vspace*{4cm}

\question[8]
For how many positive integers $b$ is $\log_b 729$ a positive integer?
\vspace*{4cm}

\newpage

\question[10]
If $\log_3(\log_3(\log_3(x))) = 2$, then how many digits are in $x$?
\vspace*{4cm}

\question[12]
Find the domain and range of $f(x) = 2\log_3(x^2 - 4x - 5)$.
\vspace*{5cm}

\newpage

\question[10]
Let $f(x) = 3x^2 - 7$ and $g(f(4)) = 9$. What is $g(f(-4))$?
\vspace*{4cm}

\question[12]
Let $f(x) = 1 - \frac{1}{x}$.

(a) Find $f(f(x))$.
\vspace*{3cm}

(b) Find $f(f(f(x)))$.
\vspace*{3cm}

(c) Find $f(f(f(f(x))))$.
\vspace*{3cm}

(d) Find $f^{34}(5)$.

\textit{Hint: Compare $f^4(x)$ to $f(x)$. Notice anything interesting? If you don't, then you should find $f^4(x)$ again.}
\vspace*{3cm}

\newpage

\question[10]
If $8^x = 27$, then what is $4^{2x - 3}$?

\textit{Hint:} Express both sides in terms of powers of 2 and 3.
\vspace*{4cm}

\question[8]
Let $f$ be a function whose graph passes through the points $(2, 3)$, $(4, 7)$, and $(8, 12)$. Suppose $f$ has an inverse. Name three points that must be on the graph of $y = f^{-1}(x)$.
\vspace*{4cm}

\question[10]
I have just won a lottery that will pay me $\$1,000,000$ in 10 years. A company offers to buy my winning ticket today for $\$300,000$.

(a) If the annually compounded interest rate is 9\%, should I take the offer?
\vspace*{3cm}

(b) For what annually compounded interest rate is my lottery ticket worth $\$300,000$ today?
\vspace*{3cm}

\newpage

\question[12]
Evaluate $\log_2 8$, $\log_2 16$, and $\log_2(8 \cdot 16)$.

(a) Evaluate $\log_3 \frac{1}{9}$, $\log_3 \sqrt{3}$, and $\log_3 \left(\frac{1}{9} \cdot \sqrt{3}\right)$.
\vspace*{3cm}

(b) Do you notice a relationship among $\log_a b$, $\log_a c$, and $\log_a(bc)$? Can you prove it?

\textit{Hint: Let $x = \log_a b$, $y = \log_a c$, and $z = \log_a(bc)$. Use exponential notation.}
\vspace*{4cm}

\question[8]
At how many points does the parabola $y = x^2$ intersect the exponential curve $y = 2^x$?
\vspace*{4cm}

\newpage

\end{questions}

\end{document} 