\documentclass[12pt]{exam}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}

% Header and Footer
\pagestyle{headandfoot}
\firstpageheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 7 of Algebra 2}
\runningheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 7 of Algebra 2}
\firstpagefooter{}{Page \thepage\ of \numpages}{}
\runningfooter{}{Page \thepage\ of \numpages}{}

% Title
\newcommand{\examtitle}{Quadratic Functions and Equations}

% Instructions
\newcommand{\instructions}{
    \noindent\rule{\textwidth}{0.5pt}
    \begin{center}
    \textbf{Instructions:} Answer all questions to the best of your ability. Show all your work in the space provided for full credit.
    \end{center}
    \noindent\rule{\textwidth}{0.5pt}
}

\begin{document}

\begin{center}
\textbf{\Large \examtitle} \\
\vspace{0.5cm}
\makebox[0.4\textwidth]{Name: \enspace\hrulefill}
\hspace{0.1\textwidth}
\makebox[0.4\textwidth]{Date: \enspace\hrulefill}
\end{center}

\instructions
\vspace{0.5cm}

\begin{questions}

\pointsinrightmargin

\question[10]
The solutions of the equation $x^2 + 2x + 2 = 0$ are $x = p$ and $x = q$. Find a quadratic equation with solutions $y = \frac{1}{p}$ and $y = \frac{1}{q}$.

\textit{Hint: First find the values of $p$ and $q$, then use the relationship between roots and coefficients.}
\vspace*{4cm}

\question[8]
Find the vertex and axis of symmetry of the parabola that is the graph of the equation $y = x^2 + 2x + 5$.
\vspace*{3cm}

\question[8]
Graph the equation $x = 2y^2 - 4y + 1$. What is its vertex?
\vspace*{4cm}

\newpage

\question[12]
Find all solutions to each of the following equations:
\begin{parts}
    \part $r^2 - 7r = 0$
    \vspace*{2cm}
    
    \part $x^2 + 3x = 7x - x^2$
    \vspace*{2cm}
    
    \part $2x^2 = 242$
    \vspace*{2cm}
    
    \part $16 - y^2 = -4$
    \vspace*{2cm}
\end{parts}

\newpage

\question[10]
Suppose $x^2 + 7bx + 10b^2 = 0$.
\begin{parts}
    \part Solve for $x$ in terms of $b$.
    \vspace*{3cm}
    
    \part Find all $b$ such that $x = 25$ is a solution of the equation.
    \vspace*{3cm}
\end{parts}

\newpage

\question[12]
Consider the quadratic $3y^2 - y - 12$.
\begin{parts}
    \part Notice that this quadratic cannot be factored into the product of two binomials with integer coefficients. Does this mean that the quadratic does not have any real roots?
    \vspace*{3cm}
    
    \part If the answer to part (a) is "no," then explain how we know that the quadratic does have real roots.
    
    \textit{Hint: Use the discriminant.}
    \vspace*{3cm}
    
    \part Suppose the quadratic has roots $y = r$ and $y = s$. Find a quadratic with roots $r + 2$ and $s + 2$.
    
    \textit{Hint: Use the quadratic formula to find the exact values of $r$ and $s$.}
    \vspace*{4cm}
\end{parts}

\newpage

\question[8]
Graph the equation $3(x + y)^2 = 6xy + 27$.

\textit{Hint: Expand and simplify the equation first.}
\vspace*{5cm}

\question[10]
Suppose the quadratic $x^2 + bx + c$ equals 0 when $x = r$ or $x = s$. If $r^2s + rs^2 = 10$, and $b$ and $c$ are integers, find all possible ordered pairs $(b, c)$.

\textit{Hint: Factor $r^2s + rs^2$.}
\vspace*{5cm}

\newpage

\question[12]
Show that the two values
$$x = \frac{2c}{-b \pm \sqrt{b^2 - 4ac}}$$
are the roots of the equation $ax^2 + bx + c = 0$, where $a$, $b$, and $c$ are constants, and $a$ and $c$ are nonzero.

\textit{Hint: That expression looks like the quadratic formula turned upside down.}

\textit{Hint: Let $x = \frac{1}{y}$ in $ax^2 + bx + c = 0$. What is $y$?}
\vspace*{6cm}

\end{questions}

\end{document} 