\documentclass[12pt]{exam}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}

% Header and Footer
\pagestyle{headandfoot}
\firstpageheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 6 of Algebra 2}
\runningheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 6 of Algebra 2}
\firstpagefooter{}{Page \thepage\ of \numpages}{}
\runningfooter{}{Page \thepage\ of \numpages}{}

% Title
\newcommand{\examtitle}{Irrational and Complex Numbers}

% Instructions
\newcommand{\instructions}{
    \noindent\rule{\textwidth}{0.5pt}
    \begin{center}
    \textbf{Instructions:} Answer all questions to the best of your ability. Show all your work in the space provided for full credit.
    \end{center}
    \noindent\rule{\textwidth}{0.5pt}
}

\begin{document}

\begin{center}
\textbf{\Large \examtitle} \\
\vspace{0.5cm}
\makebox[0.4\textwidth]{Name: \enspace\hrulefill}
\hspace{0.1\textwidth}
\makebox[0.4\textwidth]{Date: \enspace\hrulefill}
\end{center}

\instructions
\vspace{0.5cm}

\begin{questions}

\pointsinrightmargin

\question[8]
Rationalize the denominators of the following:
\begin{parts}
    \part $\frac{3}{\sqrt{6}}$
    \vspace*{2cm}
    
    \part $\frac{2}{\sqrt{9}}$
    \vspace*{2cm}
\end{parts}

\question[10]
Rationalize the denominator of $\frac{2}{\sqrt{2} - \sqrt{5} + \sqrt{7}}$.
\vspace*{4cm}

\question[12]
Simplify: $\frac{1}{\sqrt{100} + \sqrt{99}} + \frac{1}{\sqrt{99} + \sqrt{98}} + \frac{1}{\sqrt{98} + \sqrt{97}} + \cdots + \frac{1}{\sqrt{3} + \sqrt{2}} + \frac{1}{\sqrt{2} + \sqrt{1}}$
\vspace*{5cm}

\question[8]
Solve $3 - 2\sqrt{x} = 7$.
\begin{parts}
    \part $\{-2\}$
    \part $\{4\}$
    \part $\{-4\}$
    \part $\emptyset$
\end{parts}
\vspace*{3cm}

\question[10]
Simplify. If no simplification is possible, say so.
$$\sqrt[3]{24} - \sqrt[3]{56} + \sqrt[3]{81}$$
\vspace*{4cm}

\newpage

\question[12]
Find all complex numbers $a + bi$ such that $a + bi = (a + bi)^2$.

\textit{Hint: Write each side as a complex number in terms of $a$ and $b$. Build a system of equations by considering the real and imaginary parts of both sides.}
\vspace*{6cm}

\question[8]
Simplify $(i - i^{-1})^{-1}$.
\vspace*{4cm}

\question[10]
If $x = \frac{1 - i\sqrt{3}}{2}$, then what complex number is equal to $\frac{1}{x^2 - x}$?
\vspace*{5cm}

\newpage

\question[12]
For two positive numbers $x$ and $y$:
\begin{parts}
    \part The arithmetic mean of two numbers $x$ and $y$ is the number $\frac{x + y}{2}$. If $x$ and $y$ are rational numbers, what can you conclude about their arithmetic mean?
    \vspace*{3cm}
    
    \part The geometric mean of two positive numbers $x$ and $y$ is the number $\sqrt{xy}$. If $x$ and $y$ are positive rational numbers, can you conclude that their geometric mean is also rational? Explain.
    \vspace*{3cm}
\end{parts}

\question[10]
Suppose $x$ is rational and $z$ is irrational. Prove that $x + z$ is irrational.

\textit{Hint: Let $x = \frac{a}{b}$ where $a$ and $b$ are integers, and use an indirect proof by assuming that $x + z$ is a rational number $\frac{c}{d}$.}
\vspace*{5cm}

\end{questions}

\end{document} 