\documentclass[12pt]{exam}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}

% Header and Footer
\pagestyle{headandfoot}
\firstpageheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 9 of Algebra 2}
\runningheader{Integrated Algebra 2 and Precalculus}{}{Exam: Chapter 9 of Algebra 2}
\firstpagefooter{}{Page \thepage\ of \numpages}{}
\runningfooter{}{Page \thepage\ of \numpages}{}

% Title
\newcommand{\examtitle}{Systems of Linear Equations: From Two Variables to Matrices}

% Instructions
\newcommand{\instructions}{
    \noindent\rule{\textwidth}{0.5pt}
    \begin{center}
    \textbf{Instructions:} Answer all questions to the best of your ability. Show all your work in the space provided for full credit.
    \end{center}
    \noindent\rule{\textwidth}{0.5pt}
}

\begin{document}

\begin{center}
\textbf{\Large \examtitle} \\
\vspace{0.5cm}
\makebox[0.4\textwidth]{Name: \enspace\hrulefill}
\hspace{0.1\textwidth}
\makebox[0.4\textwidth]{Date: \enspace\hrulefill}
\end{center}

\instructions
\vspace{0.5cm}

\begin{questions}

\pointsinrightmargin

\question[8]
The midpoint of the segment connecting $(a, b)$ and $(b, a)$ is $(x, y)$. Express $y$ in terms of $x$.
\vspace*{4cm}

\question[10]
At what two points do the graphs of $y = -2x^2 + 5x - 3$ and $4x - y = -12$ intersect?
\vspace*{4cm}

\question[8]
At how many points does the graph of the equation $3x + y = 7$ intersect the graph of the equation $x^2 + y^2 + 6x + 10y = 2$?
\vspace*{4cm}

\newpage

\question[10]
A diameter of a circle is a segment with endpoints on the circle such that the segment passes through the center of the circle. Find the equation of the circle that has center $(3, 5)$ and a diameter with length 6 units.
\vspace*{5cm}

\question[12]
Graph the equation $x = 2y^2 - 4y + 1$. What is its vertex?
\vspace*{5cm}

\newpage

\question[10]
A parabola has its vertex at $(4, -5)$ and meets the $x$-axis at two points that are on opposite sides of the $y$-axis. Suppose this parabola is the graph of $y = ax^2 + bx + c$. Determine which of $a$, $b$, and $c$ must be positive, which must be negative, and which could be either.

\textit{Hint: What direction does the parabola open? What is the vertex of $y = ax^2 + bx + c$ in terms of $a$, $b$, and $c$?}

\textit{Hint: What does the fact that the parabola meets the $x$-axis on opposite sides of the $y$-axis tell us about the roots of $ax^2 + bx + c = 0$?}
\vspace*{5cm}

\question[8]
For how many values of $a$ is it true that the line $y = x + a$ passes through the vertex of the parabola $y = x^2 + a^2$?
\vspace*{4cm}

\newpage

\question[10]
If the parabola defined by $y = ax^2 + b$ is tangent to the line $y = x$, then calculate the constant $a$. (A line is tangent to a parabola if it touches the parabola at one point, but is otherwise always "outside" the parabola.)
\vspace*{5cm}

\question[12]
Find the vertex and axis of symmetry of the graph of $x = \frac{1}{6}(y - 4)^2 - 1$. Graph the equation.
\vspace*{5cm}

\newpage

\question[10]
Ellipse $\mathcal{E}$ has a horizontal major axis, one focus at $(4, 1)$, and one axis with an endpoint at $(6, 5)$. Find an equation whose graph is this ellipse.
\vspace*{5cm}

\question[8]
Graph each of the equations below, then find the foci, center, and lengths of the axes of each graph.

(a) $\frac{(x-3)^2}{9} + y^2 = 1$
\vspace*{4cm}

\newpage

\textbf{Extra Credit:}
$P$ is a fixed point on the diameter $\overline{AB}$ of a circle. Prove that for any chord $\overline{CD}$ of the circle that is parallel to $\overline{AB}$, we have $PC^2 + PD^2 = PA^2 + PB^2$.

\textit{Hint: Suppose the circle is the graph of $x^2 + y^2 = r^2$, and let $P$ be $(p, 0)$.}
\vspace*{5cm}

\end{questions}

\end{document} 